---
title: "HW 6.2"
author: "Maddie Drebing"
format: html
embed-resources: true
editor: visual
---

## Homework 6.2

```{r}
library(tidyverse)
library(moderndive)
theme_set(theme_minimal())
```

```{r}
data(bikes, package = "bayesrules")
glimpse(bikes)

```

## Question 0

there are 500 rows

each row represents the data of a ridership for a single day

January 2011 - December 2012

6946 rides

34 MPH

## Question 1

```{r}
get_correlation(rides ~ temp_feel,
                data = bikes) %>% round(2)
```

0.58

When the `temp_feel` goes up by one standard deviation, the rides should increase by 0.58 of a standard deviation.

```{r}
get_correlation(rides ~ windspeed,
                data = bikes) %>% round(2)
```

The correlation between number of rides and what the wind speed is -0.19.

This means when `windspeed` goes up by one standard deviation rides would go down by 0.19 of a standard deviation.

## Question 2

```{r}
bikes <- bikes %>% 
  mutate(wind_kph = windspeed * 1.61)
```

```{r}
get_correlation(windspeed ~ wind_kph,
                data = bikes)
```

The correlation between wind speed in MPH and wind speed in KPH is 1. Every KPH is 1.6x of its MPH and vice versa. It is a unit conversion.

## Question 3

```{r}
mph_model <- lm(rides ~ windspeed,
                data = bikes)
get_regression_table(mph_model) %>% 
  select(term, estimate)
```

```{r}
kph_model <- lm(rides ~ wind_kph,
                data = bikes)
get_regression_table(kph_model) %>% 
  select(term, estimate)
```

The intercepts of the two are the same since 0 MPH = 0 KPH.

the slopes are different, they are off by 1.61.

55.523/34.486 = 1.61

## Question 4 

$$
rides = 4205.065 - 34.486 (20 MPH) = 3515.345
$$

3515 riders for a 20 MPH wind day

$$
rides = 4205.065 - 55.523 (20 KPH) = 3094.605
$$

3095 riders on a 20 KPH day

## Question 5

```{r}
bikes <- bikes %>% 
  mutate(temp_c = (temp_feel-32)*(5/9))
```

```{r}
mult_model <- lm(rides ~ wind_kph + temp_c,
                 data = bikes)
get_regression_table(mult_model) %>% 
  select(term, estimate)
```

For every 1ºC increase in temperature, the model expects an increase of 143 in ridership.

For every 1 KPH increase in wind speed, the model expects a decrease of 20 in ridership.

the intercept tells us that expected ridership on a calm day with a temperature of 0ºC would be 942.

## Question 6

```{r}
pred_df <- tibble(
  situation = 1:3,
  temp_c = c(25, 15, 10),
  wind_kph = c(15, 5, 40)
)

pred_df$predicted_ridership <- predict(mult_model, newdata = pred_df)
print(pred_df)
```

there is a prediction of 4224 riders, 2991 riders, and 1581 riders for situations 1, 2, and 3.

## Question 7

```{r}
bikes <- bikes %>% mutate(weekend = as.factor(weekend))
weekend_model <- lm(rides ~ wind_kph + weekend + temp_c,
                    data = bikes)
get_regression_table(weekend_model) %>%
  select(term, estimate)
```

From the data we can see that the coefficient values for our `weekend_model`. The intercept `wind_kph` and `temp_c` coefficients fuctions in the same way in the data above. The coefficient for `weekend` says that on a weekend the model expects ridership to be 714 riders fewer than a weekday

## Question 8

```{r}
library(ggeffects)
predictions <- ggpredict(weekend_model,
                         terms = c("weekend"),
                         ci_level = NA) # don't think about confidence intervals yet (soon!)
predictions
```

While the temperature and wind speed stay at the average the expected ridership for a weekend is 2970 riders. On a weekday the expected ridership is 3683 riders.

## Question 9

```{r}
bikerides <- predict(weekend_model)
residuals <- bikes$rides - bikerides
broom::tidy(residuals)
residuals
```

the biggest residual happened on day 439 with an absolute value of 3409.346.

the model was wrong. In the original data it is 8-29-2018 when Hurricane Sandy was coming and large cities were shutting down leading to a low ridership --20
