---
title: "HW MD Ch7"
author: "Maddie Drebing"
format: html
embed-resources: true
editor: visual
---

## Ch7

```{r}
#| message: false
library(tidyverse)
library(here)
library(moderndive)
theme_set(theme_minimal())

nobel_winners <- read_csv("https://raw.githubusercontent.com/vaiseys/dav-course/main/Data/nobel_winners.csv")
```

```{r}
nobel_winners_flt <- nobel_winners %>% 
  filter(category != "Peace") %>% 
  mutate(is_us = if_else(birth_country == "United States of America", 1, 0))

true_prop <- nobel_winners_flt %>% 
  group_by(is_us) %>% 
  summarise(prop = n()/nrow(nobel_winners_flt))

ggplot() + 
  coord_cartesian(xlim = c(0,1), ylim = c(0,1)) + 
  geom_vline(xintercept = true_prop[2,2][[1]], linetype = "dashed")  + 
  labs(x = "Proportion")
```

The true proportion of US-born Nobel Prize winners is 0.306317

```{r}
set.seed(33)
# Random sample 
our_sample <- sample_n(nobel_winners_flt, 
                       size = 25, 
                       replace = F)

sample_prop <- our_sample %>% 
  group_by(is_us) %>% 
  summarise(prop = n()/nrow(our_sample))

# How does our proportion look like? 
ggplot() + 
  coord_cartesian(xlim = c(0,1), ylim = c(0,1)) + 
  geom_vline(xintercept = true_prop[2,2][[1]], linetype = "dashed")  + 
  geom_vline(xintercept = sample_prop[2,2][[1]], linetype = "dotted", col = "red") +
  labs(x = "Proportion")
```

We would be pretty far off since it is almost at 0.50

## Question 1

The population is the list of Nobel Winners

the sampling frame is `nobel_winners_flt` (scientific nobel winners)

the black line represents the population of Nobel Winners who are US-born

the red line represents the sample proportion

## Question 2

```{r}
set.seed(33)
repeated_sample <- rep_sample_n(nobel_winners_flt,
                                size = 25,
                                reps = 50)
```

```{r}
sample_proportions <- repeated_sample %>%
  group_by(replicate) %>%
  summarize(proportion = mean(is_us))

mean_sample_proportion <- mean(sample_proportions$proportion)

ggplot(sample_proportions, aes(x=proportion)) + 
  coord_cartesian(xlim = c(0,1)) + 
  geom_histogram(binwidth = 0.05, boundary = 0.4, color = "white") +
  geom_vline(xintercept = mean(nobel_winners_flt$is_us), linetype = "dashed")  + 
  geom_vline(xintercept = mean_sample_proportion, linetype = "dotted", col = "red") +
  labs(x = "Proportion of 25-winner samples that were US-born")
```

## Question 3

```{r}
set.seed(33)
repeated_sample_100 <- rep_sample_n(nobel_winners_flt,
                                size = 100,
                                reps = 50)
```

```{r}
sample_proportions_100 <- repeated_sample_100 %>%
  group_by(replicate) %>%
  summarize(proportion = mean(is_us))

mean_sample_proportion_100 <- mean(sample_proportions_100$proportion)

ggplot(sample_proportions_100, aes(x=proportion)) + 
  coord_cartesian(xlim = c(0,1)) + 
  geom_histogram(binwidth = 0.05, boundary = 0.4, color = "white") +
  geom_vline(xintercept = mean(nobel_winners_flt$is_us), linetype = "dashed")  + 
  geom_vline(xintercept = mean_sample_proportion_100, linetype = "dotted", col = "red") +
  labs(x = "Proportion of 100-winner samples that were US-born")
```

The distribution of the 25 winner sample is much more fluid than the 100 winner sample. A smaller data set will allow many different results over repetitions than a larger one.

## Question 4

```{r}
std_dev_sample_25 <- sd(sample_proportions$proportion)

std_dev_sample_100 <- sd(sample_proportions_100$proportion)


std_dev_sample_25
std_dev_sample_100
```

The sd for the 100 winner sample is smaller than the 25 winner by more than double, indicating results are jumbled much closer to the median.

## Question 5

```{r}
nobel_winners_w <- nobel_winners %>%
  filter(!is.na(gender))

women_true_prop <- nobel_winners_w %>%
  group_by(gender) %>% 
  summarise(prop = n()/nrow(nobel_winners_w))
```

the true proportion of women winners is 0.05302227

```{r}
nobel_winners_w <- nobel_winners_w %>% 
  mutate(is_w = if_else(gender == "Female", 1, 0))

```

```{r}
ggplot() + 
  coord_cartesian(xlim = c(0,1), ylim = c(0,1)) + 
  geom_vline(xintercept = mean(nobel_winners_w$is_w), linetype = "dashed")  + 
  labs(x = "Proportion")
```

```{r}
set.seed(33)
repeated_sample_w <- rep_sample_n(nobel_winners_w,
                                size = 100,
                                reps = 50)
```

```{r}
sample_proportions_w <- repeated_sample_w %>%
  group_by(replicate) %>%
  summarize(proportion = mean(is_w))

mean_sample_proportion_w <- mean(sample_proportions_w$proportion)
```

```{r}
ggplot(sample_proportions_w, aes(x=proportion)) + 
  coord_cartesian(xlim = c(0,1)) + 
  geom_histogram(binwidth = 0.02, boundary = 0.4, color = "white") +
  geom_vline(xintercept = mean(nobel_winners_w$is_w), linetype = "dashed")  + 
  geom_vline(xintercept = mean(sample_proportions_w$proportion), linetype = "dotted", col = "red") +
  labs(x = "Proportion of 100-winner samples that were woman-identifying")
```

it is nearly spot on the the true value calculated

0.0542 to our calculation of 0.05302227

## Question 6

Central Limit Theorem (CLT) is a statistical concept that states when you have a large enough sample size from a population, the distribution of the sample means will be approximately normally distributed.
