---
title: "Chapter 1-3"
format: html
embed-resources: true
editor: visual
---

## Chapter 1-3

## Question 1

Those who exercise a lot will have a lower BMI

```{r}
library(tidyverse)
# Read in the data 
exercise_data <- read_csv("https://raw.githubusercontent.com/vaiseys/dav-course/main/Data/visualize_data.csv")
glimpse(exercise_data)
```

```{r}
cor(exercise_data$Exercise, exercise_data$BMI)
```

```{r}
ggplot(data = exercise_data,
       mapping = aes(x = "Exercise", 
      y = "BMI")) +
  geom_point()
```

one point in the middle of the plot.

## Question 2

```{r}
library(causact)
glimpse(corruptDF)
```

the CPI2017 is The Corruption Perceptions Index score for 2017. A county's score is idicates the percieved lever of corruption on a scale from 0-100.

HDI2017 is The human development index score for 2017. it measures achievement in basic dimensions of human development across countries. Simple unweighted avg of a nation legevity.

## Question 3

```{r}
ggplot(data = corruptDF,
       mapping = aes(x = HDI2017,
                     y = CPI2017)) + 
  geom_point(alpha = .2)
```

The relationship between the HDI and CPI is the higher the HDI the greater the CPI

```{r}
colnames(exercise_data)[colnames(exercise_data) == "...1"] = "HDI"
colnames(exercise_data)[colnames(exercise_data) == "...2"] = "CPI"
```

## Question 4

```{r}
ggplot(data = corruptDF,
       mapping = aes(x = HDI2017,
                     y = CPI2017)) +
  geom_point() +
  labs(x = "HDI2017",
       y = "CPI2017", 
       title = "Relationship between HDI2017 and CPI2017") +
 geom_smooth(method = "gam", se = FALSE)
```

```{r}
ggplot(data = corruptDF,
       mapping = aes(x = HDI2017,
                     y = CPI2017)) +
  geom_point() +
  labs(x = "HDI2017",
       y = "CPI2017", 
       title = "Relationship between HDI2017 and CPI2017") +
 geom_smooth(method = "le", se = FALSE)
```

The gam method is more preferable and easier to read the direction of the scatter plot. It shows the relationship between the two much clearer.

## Question 5

```{r}
ggplot(data = corruptDF,
       mapping = aes(x = HDI2017,
                     y = CPI2017, fill = region, color = region)) +
  geom_point(color = "navy", fill = "grey", alpha = .5) + 
  
  labs(x = "HDI2017",
       y = "CPI2017", 
       title = "Relationship between HDI2017 and CPI2017") +
 geom_smooth(method = "gam", se = FALSE)
```

```{r}
ggplot(data = corruptDF,
       mapping = aes(x = HDI2017,
                     y = CPI2017, color = region)) +
  geom_point(color = "navy", fill = "grey", alpha = .5) + 
  
  labs(x = "HDI2017",
       y = "CPI2017", 
       title = "Relationship between HDI2017 and CPI2017") +
 geom_smooth(method = "gam", se = FALSE) +
facet_wrap(~ region, nrow = 2)
```

In the six different regions patterns are very cluttered and hard to read. Another way to see these trends would be a scatter plot used the the facet_wrap() function.

## Question 6

```{r}
ggplot(data = corruptDF,
       mapping = aes(x = CPI2017,
                     y = HDI2017, 
                     fill = region, 
                     color = region)) + 
  labs(x = "CPI2017", y = "HDI2017", 
       title = "Region - CPI2017 vs HDI2017") +
  geom_point(size = 1.5, color = "navy", fill = "grey") + 
  geom_smooth(method = "gam", se = FALSE)+ 
  scale_x_reverse()
```

## Question 7

```{r}
ggplot(data = corruptDF,
       mapping = aes(x = CPI2017,
                     y = HDI2017, 
                     fill = region, 
                     color = region)) + 
  labs(x = "CPI2017", y = "HDI2017", 
       title = "Relationships between CPI and HDI",
       subtitle = "Grouped by Region",
       caption = "The data comes from 174 observations, each observation
       being a country, where the relationship between the HDI and CPI
       exist.") +
  geom_point(size = 1.5, color = "navy", fill = "grey") + 
  geom_smooth(method = "gam", se = FALSE)+ 
  scale_x_reverse()
```

## Question 8

```{r}
ggsave("corruption_by_region.png")
```
