---
title: "DV 4 HW"
author: "Maddie Drebing"
format: html
embed-resources: true
editor: visual
---

## Chapter 4 - DV

## Question 1

```{r}
library(tidyverse)
tv_ratings <- read_csv("https://raw.githubusercontent.com/vaiseys/dav-course/main/Data/tv_ratings.csv")
glimpse(tv_ratings)
```

```{r}
tv_long <- tv_ratings %>% 
  group_by(title) %>% 
  summarise(num_seasons = n()) %>% 
  ungroup() %>% 
  left_join(tv_ratings, by = "title") 

tv_long <- tv_long %>% 
  filter(num_seasons >= 5)
```

```{r}
ggplot(data = tv_long,
       mapping = aes(x = av_rating,
                     y = seasonNumber,
                     group = title)) +
  geom_line() 
```

the less seasons the better rating. less than 5 = typically an 8+ rating.

## Question 2

```{r}
tv_long <- tv_ratings |>
  group_by(title) |>
  mutate(nseasons = max(seasonNumber)) |>
  filter(nseasons >= 5) |>
  ungroup()

ggplot(data = tv_long,
       mapping = aes(x = av_rating,
                     y = seasonNumber,
                     group = title)) +
  geom_line(alpah = .2)+
  facet_wrap(~genres) +
  theme_minimal()
```

Actions, crime, and drama have the most seasons.

As the show gets into later seasons the ratings increase.

the show that plummeted was "Are you Afraid of the Dark?"

## Question 3

```{r}
tv_ratings |>
  filter(av_rating >= 9) |>
  ggplot(tv_ratings,
         mapping = aes(x = genres)) +
  geom_bar() +
  coord_flip()
```

The coord_flip flips the x and y axis.

Drama has the most rated shows.

## Question 4

```{r}
comedies_dramas <- tv_ratings %>% 
  mutate(is_comedy = if_else(str_detect(genres, "Comedy"), 
                             1, 
                             0)) %>% # If it contains the word comedy then 1, else 0
  filter(is_comedy == 1 | genres == "Drama") %>% # Keep comedies and dramas
  mutate(genres = if_else(genres == "Drama", # Make it so that we only have those two genres
                          "Drama", 
                          "Comedy"))

glimpse(comedies_dramas)
```

```{r}
comedies_dramas |>
  ggplot(comedies_dramas,
         mapping = aes(x = av_rating,
                       fill = genres)) +
  geom_density(alpha = .4)
```

There are more higher rated comedies than there are dramas. Your prediction was correct.

## Question 5

```{r}
comedies_dramas |>
  ggplot(comedies_dramas,
         mapping = aes(x = av_rating,
                       fill = genres)) +
  geom_histogram()
```

```{r}
comedies_dramas |>
  ggplot(comedies_dramas,
         mapping = aes(x = av_rating,
                       fill = genres)) +
  geom_freqpoly()
```

The geom_freqpoly() looks like it gives you the most direct numbers. I think you can pull the most accurate information from the geom_freqpoly().

## Question 6

```{r}
comedies_dramas |>
  ggplot(comedies_dramas,
         mapping = aes(x = av_rating,
                       y = share)) +
  geom_bin_2d()
```

I see that majority of the ratings and shares are at 8 and just above 0.

in comparison to a scatter plot it shows similar information on where the points are clustered but it does not show where the most points are when overlapping.

```{r}
comedies_dramas |>
  ggplot(comedies_dramas,
         mapping = aes(x = av_rating,
                       y = share)) +
  geom_point()
```

```{r}
comedies_dramas |>
  ggplot(comedies_dramas,
         mapping = aes(x = av_rating,
                       y = share,
                       fill = genres)) +
  geom_bin_2d(alpha = .4)
```

I notice drama has a lot less shares and worse ratings compared to comedy that has more shares and higher ratings.

The outlier is the drama Dekalog.
